---
title: "MSDS Case Study 01, SMU"
author: "M Nepal"
date: "June 29, 2017"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The case study is based on **Beers** and **Brewery** dataset provided in <a href="https://github.com/m-nepal/MSDS6306_CaseStudy01/tree/master/Data"> Data Directory</a> in the project. 
### Dataset descriptions:

##### Beers
* Name: Name of the beer.
* Beer_ID: Unique identier of the beer.
* ABV: Alcohol by volume of the beer.
* IBU: International Bitterness Units of the beer.
* Brewery_ID: Brewery id associated with the beer.
* Style: Style of the beer.
* Ounces: Ounces of beer.

##### Breweries
* Brew_ID: Unique identier of the brewery.
* Name: Name of the brewery.
* City: City where the brewery is located.
* State: State where the brewery is located.

## Required Library
Download <a href="https://cran.r-project.org/">R<a/> and <a href="https://www.rstudio.com/products/rstudio/download/">Rstudio</a> and install as you follow the code.

Once you download and install R and R studio, you need following libraries to complete the analyisis.
<ul><li><a href="http://ggplot2.org/">ggplot2<a/></li><li><a href="https://cran.r-project.org/web/packages/dplyr/index.html">dplyr<a/></li></ul>
**SessionInfo :**
R version 3.3.2 (2016-10-31)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows >= 8 x64 (build 9200)`

## What we are trying to achieve?
This project analyses the data with some of the basic features available in **R**.There are plenty of other analyis we can conduct in <a href="https://cran.r-project.org/">R<a/>, but this project focuses on the following:
<ul>
<li>Numbers of breweries per state</li>
<li>Merging <a href="https://github.com/m-nepal/MSDS6306_CaseStudy01/tree/master/Data">**Beers** and **Brewery**<a/> data set based on common brewery_id</li>
<li>Reporting number's of NA's in each column</li>
<li>Computing median of ABV and IBU per state and plotting bar chart</li>
<li>Reporting number's of NA's in each column</li>
<li>Computing maximum alcohalic beer and most bitter beer by state</li>
<li>Finding the relation between bitterness and it's alcohalic content</li>
</ul>

### Set up
```{r}
# Clean up all the variables
rm(list=ls())
ls()
beerDir <- "C:/Users/mohan/Desktop/MSDS6306_CaseStudy01/Data/Beers.csv"
breweryDir <- "C:/Users/mohan/Desktop/MSDS6306_CaseStudy01/Data/Breweries.csv"
#install following package and load the library simply uncommenting following code
#install.packages("ggplot2")
#install.packages("dplyr")
#library(dplyr)
#library(ggplot2)
```

### Analysis for data sets
```{r}
# importing data set to variables 
beers <- read.csv(beerDir, sep = ",", header = TRUE,na.strings = c("","NA"))
breweries <- read.csv(breweryDir, header = TRUE, sep=",",na.strings =c("","NA"))
# check the content of beers and breweries
head(beers)
head(breweries)
```
#### Find numbers of breweries per state
```{r}
# There are different ways to find out the number of brewery per state but in this analysis following method is used
brewery_ps_way1 <- data.frame(table(breweries$State))
brewery_ps_way2 <- aggregate(rep(1, length(breweries$State)),by=list(breweries$State), sum)
# check the number of brwery per state by calling head method
head(brewery_ps_way1)
# In this result, Var1 represents the State and  Freq represents the number of brewery in that state
head(brewery_ps_way2)
#in this method Group.1 column is State and x represnts the number of brewery
# Identyfying first 6 rows,  we can verify both of the method are giving the same result.
# Here dplyr package is used to give the meaningful name to coulmns header
# For presentation uncomment the library
library(dplyr)
breweriesPerState <- rename(brewery_ps_way1,State = Var1,Breweries_Count = Freq)
head(breweriesPerState)
# Full list 
breweriesPerState
```
#### Merge the data set by brewery_id
```{r}
merged_data <- merge(beers, breweries, by.x = "Brewery_id",by.y = "Brew_ID")
dim(merged_data)
# Here Beer name is listed as Name.x and Brewery name as Name.y 
# Use rename method to give some meaning full name on merged data
# use rename method from "dplyr" package
merged_new <- rename(merged_data, Beer = Name.x , Brewery = Name.y)
#printing first and last 6 obs. to check the content of  merged data
head(merged_new)
tail(merged_new)
```
#### Is there's any NA's?
```{r}
# There are different techniques that can be applied to findout missing NA's
# We can avoid blank cell to make NA's when we import as above code or we can use following code to report NA's

missing.beer <- apply(beers,2, function(x){sum(is.na(x))})
missing.brewery <- apply(breweries,2, function(x){sum(is.na(x))})
missing.merged <- apply(merged_new, 2, function(x){ sum(is.na(x))})
# another way to report missing merge
merged_2 <- replace(merged_new, merged_new == "", NA)
missing.merged_2 <- apply(merged_2, 2, function(x){ sum(is.na(x))})
# check the number's of NA's  
missing.beer
missing.brewery
missing.merged
missing.merged_2
# The NA's in beer's data set and merged data set looks similar, so we can verify the NA's numbers in data sets.
```
#### Computing median alcohal content and International bitterness unit for each state
```{r}
# Assign a new merged data set to a new variable for easy use for further analysis
cleanedData <- merged_2
# Check first 6 and last 6 rows just to make sure you have right data
head(cleanedData)
tail(cleanedData)

#Finding median ABV and IBU per state
median_ABV_IBU <- aggregate(cbind(cleanedData$ABV,cleanedData$IBU)~State, FUN = median,cleanedData,na.rm = TRUE)
head(median_ABV_IBU)
#here ABV is named as V1 and IBU as V2, so using colnames we can rename to give meaningful name
colnames(median_ABV_IBU) <- c("State","Median_ABV","Median_IBU")
# After renaming lets' check the column header once again
head(median_ABV_IBU)
```
Data can make more sense if we see the pattern in some visual interface.So let's plot the barchart for median ABV and IBU per state
##### Bar plot for ABV and IBU per state
```{r}
library(ggplot2)
ggplot(median_ABV_IBU,aes(median_ABV_IBU$State, y=median_ABV_IBU$Median_ABV))+
    geom_bar(stat="identity")+labs(x="State",y="ABV") + ggtitle("ABV median per State")

ggplot(median_ABV_IBU,aes(median_ABV_IBU$State, y=median_ABV_IBU$Median_IBU))+
  geom_bar(stat="identity")+labs(x="State",y="IBU")+ggtitle("IBU median per State")
```

#### What we see from the bar chart?
Comparing two bar graph, if the ABV content in the beer is increased the IBU also seems to be increasing. Most of the state has bitter beer with  ABV atleast 0.05. To summarise the bar graph, Higher the alcohalic content more the bitterness in the beer.

#### Which state has the MAX alcohalic beer and MOST bitter beer?
```{r}
max_ABV <- max(cleanedData$ABV,na.rm = TRUE)
max_ABV_2 <- arrange(cleanedData,desc(cleanedData$ABV))
max_ABV_3 <- head(max_ABV_2,1)
max_ABV_State <- data.frame(max_ABV_3$State,max_ABV_3$ABV)

# IBU
max_IBU <- max(cleanedData$IBU,na.rm = TRUE)
max_IBU_2 <- arrange(cleanedData,desc(cleanedData$IBU))
max_IBU_3 <- head(max_IBU_2,1)
max_IBU_State <- data.frame(max_IBU_3$State,max_IBU_3$IBU)

# let's find out which state has  the max ABV and Most bitter beer
max_ABV_State 
max_IBU_State
# CO has the maximum alcohalic beer and OR has the most bitter beer.
```
#### Summary statistics for ABV
```{r}
# summary statistics of ABV
summary(cleanedData$ABV)
```

#### Did you find the relationship between the bitterness of the beer and it's alcohalic content?
Let's plot the scatter plot to do so.
```{r}
ggplot(beers, aes(x = ABV, y = IBU)) + geom_point(na.rm=TRUE)+geom_smooth(method=lm,se=FALSE, na.rm=TRUE)

```

Athough there are outliers, there is positive correlation between two variables. More ABV content in beer leads more bitterness beer.

#### Conclusion
Analysing the provided data, Colorado has the most alcohalic beer and Orlando has most bitter beer. There were some NA's in the data sets.Orlando enjoys most bitter beer and  Colorado enjoys most alcohalic beer.And the scatter plot shows the positive correlation between ABV and IBU contents of beer.
The purpose of this project is to provide a hands on knowledge on some of the techniques like Data exploration, data wrangling, graphics using R. And I would say, the project has accomplished it's  purpose of it.
